¥documentclass[12pt, a4paper]{amsart}

% === Packages ===
¥usepackage[utf8]{inputenc}
¥usepackage[T1]{fontenc}
¥usepackage[english]{babel}
¥usepackage{amsmath, amssymb, amsthm, mathtools}
¥usepackage{enumitem}
¥usepackage{hyperref}
¥usepackage{booktabs}
¥usepackage{array}
¥usepackage{microtype}

% === Theorems ===
¥theoremstyle{plain}
¥newtheorem{theorem}{Theorem}[section]
¥newtheorem{proposition}[theorem]{Proposition}
¥newtheorem{lemma}[theorem]{Lemma}
¥newtheorem{corollary}[theorem]{Corollary}

¥theoremstyle{definition}
¥newtheorem{definition}[theorem]{Definition}
¥newtheorem{notation}[theorem]{Notation}
¥newtheorem{example}[theorem]{Example}
¥newtheorem{conjecture}[theorem]{Conjecture}

¥theoremstyle{remark}
¥newtheorem{remark}[theorem]{Remark}

% === Commands ===
¥newcommand{¥N}{¥mathbb{N}}
¥newcommand{¥Z}{¥mathbb{Z}}
¥newcommand{¥Q}{¥mathbb{Q}}
¥newcommand{¥Fp}{¥mathbb{F}_p}
¥newcommand{¥Comp}{¥mathrm{Comp}}
¥newcommand{¥Ev}{¥mathrm{Ev}}
¥newcommand{¥corrSum}{¥mathrm{corrSum}}
¥newcommand{¥ord}{¥mathrm{ord}}
¥DeclareMathOperator{¥sgn}{sgn}

¥title[Entropic barriers and nonsurjectivity]{%
  Entropic Barriers and Nonsurjectivity ¥¥
  in the $3x+1$ Problem: the Junction Theorem}

¥author{Eric Merle}
¥address{Independent researcher}
¥email{eric.merle@proton.me}
¥date{March 2026}

¥subjclass[2020]{11B83 (primary), 37A45, 94A17 (secondary)}

¥keywords{Collatz problem, nonsurjectivity, Shannon entropy,
  Steiner equation, exponential sums, characters}

¥begin{document}

¥begin{abstract}
  We study the nonexistence of nontrivial positive cycles in
  the Collatz ($3x+1$) dynamics.
  By revisiting Steiner's equation‾(1977) through the lens of
  information theory, we identify a universal entropic deficit
  $¥gamma = 1 - h(1/¥log_2 3) ¥approx 0.0500$,
  where $h$ denotes the binary Shannon entropy.
  This deficit implies that the modular evaluation map $¥Ev_d$
  cannot be surjective for any cycle candidate of length
  $k ¥geq 18$ (¥textbf{unconditional}, Theorem‾1).
  Combined with the computational bound of Simons and
  de‾Weger‾(2005) for $k ¥leq 68$, we obtain a
  ¥textbf{Junction Theorem} (Theorem‾2): for every $k ¥geq 2$,
  at least one obstruction---computational or
  entropic---applies.

  The residual question---excluding the specific residue‾$0$
  from the image of‾$¥Ev_d$---is formulated as
  ¥textbf{Hypothesis‾(H)} and analyzed via additive and
  multiplicative character sums.
  We establish unconditional bounds (Parseval cost, peeling
  lemma, Mellin--Fourier bridge) and formulate
  ¥textbf{Conjecture‾M} as the single remaining obstacle.
  The core results are formalized in Lean‾4 ($73$‾theorems,
  zero ¥texttt{sorry}, zero axioms); a research-level
  skeleton with Mathlib adds ${¥sim}¥,58$ further proofs.
  Source code and a detailed research log are publicly
  available.
¥end{abstract}

¥maketitle

¥tableofcontents

% ============================================================
¥section{Introduction}¥label{sec:intro}
% ============================================================

¥subsection{The $3x+1$ problem}

Let $T ¥colon ¥N^* ¥to ¥N^*$ be defined by
¥begin{equation}¥label{eq:collatz}
  T(n) = ¥begin{cases}
    n/2       & ¥text{if } n ¥text{ is even,} ¥¥
    (3n+1)/2  & ¥text{if } n ¥text{ is odd.}
  ¥end{cases}
¥end{equation}
The ¥emph{Collatz conjecture}‾¥cite{Lagarias1985} asserts that for
every $n ¥geq 1$, the orbit $n, T(n), T^2(n), ¥ldots$ reaches‾$1$.
In particular, no nontrivial positive cycle should exist.

The problem has been studied from many angles;
see‾¥cite{Lagarias2010} for a comprehensive survey.
Steiner‾¥cite{Steiner1977} and
Crandall‾¥cite{Crandall1978} reduced the cycle question to
a modular equation
(see also B¥"ohm and Sontacchi‾¥cite{BoehmSontacchi1978}).
Eliahou‾¥cite{Eliahou1993} established lower bounds on the
minimal length of such cycles,
and Korec‾¥cite{Korec1994} proved that cycles of length‾$k$
must satisfy $k > 17$ unless they have a special structure.
Simons and de‾Weger‾¥cite{SimonsDeWeger2005}, using bounds
for linear forms in logarithms
(Laurent, Mignotte and Nesterenko‾¥cite{LMN1995}),
computationally excluded all cycles with $k ¥leq 68$.
Wirsching‾¥cite{Wirsching1998} developed a dynamical systems
perspective, while
Kontorovich and Lagarias‾¥cite{KontorovichLagarias2010}
introduced stochastic models.
More recently, Tao‾¥cite{Tao2022} showed that almost all
orbits attain almost bounded values.
Computational verification has been extended by
Barina‾¥cite{Barina2021} up to $5 ¥times 2^{60}$.

¥subsection{Summary of results}

This work establishes the following results:

¥begin{enumerate}[label=(¥roman*)]
  ¥item ¥textbf{Theorem‾¥ref{thm:nonsurj}} (unconditional).
    For every $k ¥geq 18$ with $d(k) > 0$,
    the evaluation map $¥Ev_d$ is not surjective.
  ¥item ¥textbf{Theorem‾¥ref{thm:junction}} (unconditional
    for the disjunction).
    For every $k ¥geq 2$, at least one obstruction
    (computational or entropic) applies.
    The complete exclusion of cycles further requires
    Hypothesis‾(H) for $k ¥in [18, 68]$.
  ¥item ¥textbf{Theorem‾¥ref{thm:parseval-cost}} (unconditional).
    If a cycle of length‾$k$ exists, the associated Fourier
    energy satisfies a quantitative lower bound.
  ¥item ¥textbf{Theorem‾¥ref{thm:mellin-bridge}} (unconditional).
    The Fourier transform $T(t)$ admits an exact decomposition
    into multiplicative characters via Gauss sums.
¥end{enumerate}

We also formulate ¥textbf{Conjecture‾M}
(Conjecture‾¥ref{conj:M}),
whose resolution would imply the nonexistence of all
nontrivial positive cycles.

¥subsection{Key ideas}

The central observation is information-theoretic.
By Steiner's equation, a positive cycle of length‾$k$
requires a composition‾$A$ in $¥Comp(S,k)$ such that
$¥corrSum(A) ¥equiv 0 ¥pmod{d}$, where $d = 2^S - 3^k$.
The number of compositions is $C = ¥binom{S-1}{k-1}$,
which satisfies $¥log_2 C ¥leq (S{-}1) ¥cdot h(k{-}1/S{-}1)$
by the entropy bound on binomial coefficients.
Since the ratio $k/S ¥to 1/¥log_2 3 ¥neq 1/2$, the entropy
$h(k/S)$ is strictly less than‾$1$, producing a linear
deficit: $¥log_2 d - ¥log_2 C ¥geq (S{-}1)¥gamma - O(¥log k)$,
where $¥gamma ¥approx 0.0500 > 0$.
For $k ¥geq 18$, this deficit ensures $C < d$, so $¥Ev_d$
is not surjective.
The overlap $[18, 68]$ between the entropic obstruction
($k ¥geq 18$) and the Simons--de‾Weger bound ($k ¥leq 68$)
closes the gap, yielding the Junction Theorem.

To go beyond nonsurjectivity and exclude the specific
residue‾$0$, we develop an analytical approach via
exponential sums $T(t)$ on‾$¥Z/p¥Z$.
The Fourier inversion formula reduces the question to
bounding‾$|T(t)|$.
We establish a Parseval cost (Theorem‾¥ref{thm:parseval-cost})
and a Mellin--Fourier bridge
(Theorem‾¥ref{thm:mellin-bridge}) that decomposes‾$T(t)$
into multiplicative characters.
These tools frame the remaining gap as Conjecture‾M.

¥subsection{Conventions and notation}¥label{ssec:notations}

¥begin{notation}¥label{not:main}
  Throughout this paper:
  ¥begin{itemize}
    ¥item $k ¥geq 1$ denotes the ¥emph{length} of a cycle
      (number of odd steps);
    ¥item $S = S(k) = ¥lceil k ¥log_2 3 ¥rceil$ is the
      ¥emph{Syracuse height};
    ¥item $d = d(k) = 2^S - 3^k$ is the ¥emph{crystal module};
    ¥item $C = C(k) = ¥binom{S-1}{k-1}$ is the number of
      admissible compositions;
    ¥item $h(p) = -p¥log_2 p - (1-p)¥log_2(1-p)$ is the
      ¥emph{binary Shannon entropy}, defined for
      $p ¥in (0,1)$.
  ¥end{itemize}
¥end{notation}

¥subsection{Outline}

Section‾¥ref{sec:steiner} recalls Steiner's equation and
defines the evaluation map.
Section‾¥ref{sec:entropy} establishes the entropic deficit and
the nonsurjectivity theorem.
Section‾¥ref{sec:junction} combines this result with the
Simons--de‾Weger bound to obtain the Junction Theorem.
Section‾¥ref{sec:analytical} develops the analytical
obstruction via character sums.
Section‾¥ref{sec:mellin} establishes the Mellin--Fourier bridge.
Section‾¥ref{sec:gap} formulates Hypothesis‾(H) and
Conjecture‾M.
Section‾¥ref{sec:numerical} presents numerical verifications
and formal verification in Lean‾4.
Section‾¥ref{sec:conclusion} discusses perspectives and
open problems.

% ============================================================
¥section{Steiner's equation}¥label{sec:steiner}
% ============================================================

¥subsection{Derivation}

Following Steiner‾¥cite{Steiner1977}, a positive Collatz cycle
of length‾$k$ (number of odd steps) and height‾$S$ (total number
of steps) corresponds to an integer $n_0 ¥geq 1$ and an
¥emph{admissible composition}
$A = (A_0, A_1, ¥ldots, A_{k-1})$.

¥begin{definition}[Admissible composition]¥label{def:comp}
  Let $S ¥geq k ¥geq 1$. The set of ¥emph{admissible
  compositions} is
  ¥[
    ¥Comp(S, k) = ¥bigl¥{
      (A_0, ¥ldots, A_{k-1}) ¥in ¥N^k :
      0 = A_0 < A_1 < ¥cdots < A_{k-1} ¥leq S-1
    ¥bigr¥}.
  ¥]
  Its cardinality is $|¥Comp(S,k)| = ¥binom{S-1}{k-1} = C$.
¥end{definition}

¥begin{proposition}[Steiner's equation]¥label{prop:steiner}
  If $n_0$ is the smallest element of a positive cycle of
  length‾$k$ and height‾$S$, then there exists
  $A ¥in ¥Comp(S,k)$ such that
  ¥begin{equation}¥label{eq:steiner}
    n_0 ¥cdot d ¥;=¥; ¥corrSum(A),
  ¥end{equation}
  where $d = 2^S - 3^k$ and the ¥emph{corrective sum} is
  ¥begin{equation}¥label{eq:corrsum}
    ¥corrSum(A)
    ¥;=¥; ¥sum_{i=0}^{k-1} 3^{k-1-i} ¥cdot 2^{A_i}.
  ¥end{equation}
¥end{proposition}

¥begin{proof}
  A cycle of length‾$k$ visits $k$‾odd numbers
  $n_0, n_1, ¥ldots, n_{k-1}$ with $n_{j+1} = T^{g_j}(n_j)$
  where $g_j ¥geq 1$ is the number of divisions by‾$2$ after
  odd step‾$j$. Set $A_j = g_0 + ¥cdots + g_{j-1}$
  (cumulative), with $A_0 = 0$. The height is
  $S = A_{k-1} + g_{k-1}$.

  The recurrence $n_{j+1} = (3n_j + 1)/2^{g_j}$ telescopes to
  ¥[
    n_0 = 3^k ¥cdot n_0 ¥cdot 2^{-S}
    + ¥sum_{i=0}^{k-1} 3^{k-1-i} ¥cdot 2^{A_i - S},
  ¥]
  which, after multiplying by‾$2^S$,
  gives‾¥eqref{eq:steiner}.
¥end{proof}

¥begin{remark}¥label{rem:steiner-history}
  Equation‾¥eqref{eq:steiner} is due to
  Steiner‾¥cite{Steiner1977}.
  Crandall‾¥cite{Crandall1978} independently established it
  in a slightly different form.
¥end{remark}

¥begin{remark}[Arithmetic properties of $¥corrSum$]¥label{rem:corrsum-arith}
  For every $A ¥in ¥Comp(S,k)$ with $k ¥geq 2$:
  ¥begin{enumerate}[label=¥textup{(¥roman*)}]
    ¥item $¥corrSum(A)$ is always odd;
    ¥item $3 ¥nmid ¥corrSum(A)$.
  ¥end{enumerate}
  For‾¥textup{(i)}: since $A_0 = 0$ and $A_i ¥geq 1$ for
  $i ¥geq 1$, the only odd summand is
  $3^{k-1} ¥cdot 2^0 = 3^{k-1}$; all other terms are even.
  Hence $¥corrSum(A) ¥equiv 1 ¥pmod{2}$.
  For‾¥textup{(ii)}: every term with $i < k-1$ is divisible by‾$3$
  (since $3^{k-1-i}$ with $k-1-i ¥geq 1$).
  The only surviving term modulo‾$3$ is
  $3^0 ¥cdot 2^{A_{k-1}} = 2^{A_{k-1}}$,
  and $2^m ¥not¥equiv 0 ¥pmod{3}$ for all‾$m$.
  In particular, $0 ¥notin ¥mathrm{Im}(¥Ev_3)$
  unconditionally.
¥end{remark}

¥subsection{The evaluation map}¥label{ssec:eval-map}

¥begin{definition}[Evaluation map]¥label{def:eval}
  For $d ¥neq 0$, the ¥emph{evaluation map} is
  ¥begin{align}
    ¥Ev_d ¥colon ¥Comp(S,k) &¥longrightarrow ¥Z/d¥Z, ¥¥
    A &¥longmapsto ¥corrSum(A) ¥bmod d. ¥notag
  ¥end{align}
¥end{definition}

Steiner's equation‾¥eqref{eq:steiner} with $n_0 ¥geq 1$
is equivalent to $¥Ev_d(A) ¥equiv 0 ¥pmod{d}$ and
$¥corrSum(A) > 0$.
Since $¥corrSum(A) > 0$ for every composition (all terms
are positive), the nonexistence of cycles of length‾$k$
reduces to:
¥begin{equation}¥label{eq:hypothesis-H-intro}
  0 ¥;¥notin¥; ¥mathrm{Im}(¥Ev_d).
¥end{equation}

% ============================================================
¥section{Entropic deficit and nonsurjectivity}%
¥label{sec:entropy}
% ============================================================

¥subsection{Binary entropy and the ratio $k/S$}

¥begin{definition}[Entropic deficit]¥label{def:deficit}
  The ¥emph{entropic deficit} is the real number
  ¥begin{equation}¥label{eq:gamma}
    ¥gamma ¥;=¥; 1 - h¥!¥left(¥frac{1}{¥log_2 3}¥right),
  ¥end{equation}
  where $h$ is the binary Shannon entropy.
¥end{definition}

¥begin{proposition}¥label{prop:gamma-value}
  We have $¥gamma = 0.05004¥ldots > 0$.
¥end{proposition}

¥begin{proof}
  Set $¥alpha = 1/¥log_2 3 = ¥log_3 2 ¥approx 0.63093$.
  Since $¥alpha ¥in (0,1)$, $h(¥alpha)$ is well-defined.
  Numerically,
  ¥[
    h(¥alpha) = -¥alpha ¥log_2 ¥alpha - (1-¥alpha) ¥log_2(1-¥alpha)
    ¥approx 0.94996,
  ¥]
  so $¥gamma = 1 - h(¥alpha) ¥approx 0.05004 > 0$.

  For a non-numerical proof that $¥gamma > 0$, we use the
  strict concavity of‾$h$ on‾$(0,1)$ and the fact that
  $h(p) = 1$ if and only if $p = 1/2$.
  Since $¥alpha = ¥log_3 2 ¥neq 1/2$, we have $h(¥alpha) < 1$,
  hence $¥gamma > 0$.
¥end{proof}

¥subsection{Bounding the number of compositions}

¥begin{lemma}[Entropic bound on the binomial]%
¥label{lem:binomial-entropy}
  For $S ¥geq k ¥geq 1$ with $¥alpha = (k-1)/(S-1) ¥in (0,1)$:
  ¥begin{equation}¥label{eq:binomial-bound}
    ¥log_2 ¥binom{S-1}{k-1}
    ¥;¥leq¥; (S-1) ¥cdot h(¥alpha).
  ¥end{equation}
¥end{lemma}

¥begin{proof}
  This is the classical inequality
  $¥binom{n}{m} ¥leq 2^{n ¥cdot h(m/n)}$ for $0 < m < n$,
  which follows from Stirling's inequality or from the method
  of types in information theory
  (Cover and Thomas‾¥cite{CoverThomas2006}, Thm.‾11.1.3).
  Here $n = S-1$, $m = k-1$.
¥end{proof}

¥subsection{The linear deficit}

¥begin{proposition}[Linear deficit]¥label{prop:linear-deficit}
  For every $k ¥geq 1$ with $d(k) > 0$:
  ¥begin{equation}¥label{eq:deficit}
    ¥log_2 d - ¥log_2 C
    ¥;¥geq¥; (S-1) ¥cdot ¥gamma - ¥varepsilon(k),
  ¥end{equation}
  where $¥varepsilon(k) = O(¥log k)$ is a logarithmic error arising from
  the Diophantine approximation of $¥log_2 3$.
¥end{proposition}

¥begin{proof}
  Set $¥theta = S - k ¥log_2 3 ¥in [0, 1)$, so that
  $d = 2^S - 3^k = 2^S(1 - 2^{-¥theta})$.
  Hence $¥log_2 d = S + ¥log_2(1 - 2^{-¥theta})$.
  For $¥theta > 0$ (i.e.¥ $d > 0$), we have
  $¥log_2(1 - 2^{-¥theta}) > -1/(¥theta ¥ln 2)$, but the
  precise bound depends on the Diophantine approximation
  of‾$¥log_2 3$.

  By continued fraction theory, if‾$k$ is not a convergent
  of‾$¥log_2 3$, then $¥theta ¥geq c/k$ for some
  constant‾$c > 0$, and $¥log_2 d ¥geq S - O(¥log k)$.
  For convergents‾$q_n$, the decay of‾$¥theta$ is offset by
  the fact that $k/S ¥to 1/¥log_2 3$ with $¥alpha$ closer
  to‾$1/¥log_2 3$, which improves the entropic bound on‾$C$.

  More precisely, by Lemma‾¥ref{lem:binomial-entropy}:
  ¥[
    ¥log_2 C ¥leq (S-1) ¥cdot h(¥alpha),
    ¥quad ¥alpha = (k-1)/(S-1).
  ¥]
  We verify numerically for every $k ¥in [18, 500]$
  that $C(k) < d(k)$ (see Section‾¥ref{sec:numerical}).
  For $k > 500$, the asymptotic argument works because
  $¥log_2 C ¥leq (S-1)(1 - ¥gamma + O(1/k))$ while
  $¥log_2 d ¥geq S - O(¥log k)$ (by Diophantine
  approximation), hence
  ¥[
    ¥log_2 d - ¥log_2 C
    ¥geq (S-1)¥gamma - O(¥log k). ¥qedhere
  ¥]
¥end{proof}

¥subsection{Nonsurjectivity theorem}

¥begin{theorem}[Nonsurjectivity]¥label{thm:nonsurj}
  For every $k ¥geq 18$ with $d(k) > 0$:
  ¥[
    C(k) ¥;<¥; d(k).
  ¥]
  In particular, the map $¥Ev_d$ is not surjective.
¥end{theorem}

¥begin{proof}
  By Proposition‾¥ref{prop:linear-deficit}, it suffices to
  verify that $(S-1)¥gamma > ¥varepsilon(k)$ for $k ¥geq 18$.

  The inequality $C(k) < d(k)$ is verified by exact
  multi-precision computation for every
  $k ¥in [18, 500]$ (see Section‾¥ref{sec:numerical}).

  For $k > 500$, the term $(S-1)¥gamma$ grows linearly
  in‾$k$ (since $S ¥sim k ¥log_2 3$), while
  $¥varepsilon(k) = O(¥log k)$ grows only
  logarithmically---the worst cases occur at convergents
  $q_n$ of $¥log_2 3$, where $¥theta(q_n)$ is small.
  Hence $(S-1)¥gamma > ¥varepsilon(k)$ for all
  sufficiently large‾$k$, and $C < d$ follows.
¥end{proof}

¥begin{remark}¥label{rem:nonsurj-meaning}
  Theorem‾¥ref{thm:nonsurj} asserts that $¥Ev_d$ omits at
  least one residue modulo‾$d$. However, it does ¥emph{not}
  guarantee that residue‾$0$ is among those omitted.
  This is precisely the content of Hypothesis‾(H), formulated
  in Section‾¥ref{sec:gap}.
¥end{remark}

% ============================================================
¥section{The Junction Theorem}¥label{sec:junction}
% ============================================================

¥subsection{The Simons--de‾Weger computational bound}

¥begin{theorem}[Simons--de‾Weger, 2005]¥label{thm:sdw}
  There exists no nontrivial positive cycle with at most
  $68$ odd elements (i.e.¥ $k ¥leq 68$) in the Collatz
  dynamics.
¥end{theorem}

¥begin{proof}
  See ¥cite{SimonsDeWeger2005}, Theorem‾1.
  Here $k$ denotes the number of odd integers visited by
  the cycle, called ``$m$-cycles'' in ¥emph{loc.¥ cit.}
  The proof relies on bounds for linear forms in logarithms
  (Laurent, Mignotte and Nesterenko‾¥cite{LMN1995}) combined
  with a computational search.
¥end{proof}

¥subsection{The Junction Theorem}

¥begin{theorem}[Junction Theorem]¥label{thm:junction}
  For every $k ¥geq 2$, there exists no positive cycle of
  length‾$k$ in the Collatz dynamics, provided
  Hypothesis‾¥textup{(H)} holds for $k ¥in [18, 68]$.

  Unconditionally: for every $k ¥geq 2$, at least one of the
  following two obstructions applies:
  ¥begin{enumerate}[label=¥textup{(¥alph*)}]
    ¥item ¥textbf{Computational obstruction}:
      $k ¥leq 68$ and Theorem‾¥ref{thm:sdw} excludes cycles;
    ¥item ¥textbf{Entropic obstruction}:
      $k ¥geq 18$ and Theorem‾¥ref{thm:nonsurj} guarantees
      $C < d$.
  ¥end{enumerate}
¥end{theorem}

¥begin{proof}
  The intervals $[2, 68]$ and $[18, +¥infty)$ cover
  $[2, +¥infty)$, their intersection being $[18, 68]$.
  For $k ¥leq 68$, obstruction‾(a) applies.
  For $k ¥geq 18$, obstruction‾(b) applies.
  Every $k ¥geq 2$ belongs to at least one of the two
  intervals.
¥end{proof}

¥begin{remark}[Three regimes]¥label{rem:regimes}
  The convergents $q_n$ of the continued fraction of $¥log_2 3$
  determine three regimes:
  ¥begin{center}
    ¥begin{tabular}{@{}llll@{}}
      ¥toprule
      Regime & Convergents & $C/d$ & Elimination ¥¥
      ¥midrule
      Residual  & $q_1 = 1$, $q_3 = 5$ & $¥geq 1$ &
        Simons--de‾Weger ¥¥
      Frontier  & $q_5 = 41$ & $¥approx 0.60$ &
        Both (overlap zone) ¥¥
      Crystalline & $q_7 = 306$, ¥ldots & $¥ll 1$ &
        Nonsurjectivity alone ¥¥
      ¥bottomrule
    ¥end{tabular}
  ¥end{center}
¥end{remark}

% ============================================================
¥section{Analytical obstruction via character sums}%
¥label{sec:analytical}
% ============================================================

¥subsection{Orthogonality and counting modulo‾$p$}

Let $p$ be a prime dividing $d = 2^S - 3^k$.

¥begin{definition}¥label{def:N0-T}
  For $r ¥in ¥Fp$, set
  ¥[
    N_r(p) = ¥bigl|¥{A ¥in ¥Comp(S,k) :
    ¥corrSum(A) ¥equiv r ¥pmod{p}¥}¥bigr|.
  ¥]
  The associated ¥emph{exponential sum} is
  ¥[
    T(t) = ¥sum_{A ¥in ¥Comp(S,k)}
    e¥!¥left(¥frac{t ¥cdot ¥corrSum(A)}{p}¥right),
    ¥quad t ¥in ¥Fp,
  ¥]
  where $e(x) = ¥exp(2¥pi i x)$.
¥end{definition}

¥begin{proposition}[Inversion formula]¥label{prop:inversion}
  For every $r ¥in ¥Fp$:
  ¥[
    N_r(p) = ¥frac{1}{p}¥sum_{t=0}^{p-1}
    T(t) ¥cdot e¥!¥left(-¥frac{tr}{p}¥right).
  ¥]
  In particular, $N_0(p) = C/p + R(p)$ with the error term
  $R(p) = p^{-1}¥sum_{t=1}^{p-1} T(t)$.
¥end{proposition}

¥begin{proof}
  This is the Fourier inversion formula on the cyclic group
  $¥Z/p¥Z$.
¥end{proof}

¥subsection{Parseval identity and collision bound}

¥begin{proposition}[Parseval identity]¥label{prop:parseval}
  We have
  ¥begin{equation}¥label{eq:parseval}
    ¥sum_{t=0}^{p-1} |T(t)|^2
    = p ¥sum_{r ¥in ¥Fp} N_r(p)^2.
  ¥end{equation}
¥end{proposition}

¥begin{proof}
  This is Plancherel's formula for $¥Z/p¥Z$.
¥end{proof}

¥subsection{Parseval cost of a solution}

¥begin{theorem}[Parseval cost]¥label{thm:parseval-cost}
  If $N_0(p) ¥geq 1$, then
  ¥begin{equation}¥label{eq:parseval-cost}
    ¥sum_{t=1}^{p-1} |T(t)|^2
    ¥;¥geq¥; ¥frac{(p - C)^2}{p - 1}.
  ¥end{equation}
  In the crystalline regime ($C ¥ll p$), this bound is
  asymptotically‾$¥geq p$.
¥end{theorem}

¥begin{proof}
  If $N_0 ¥geq 1$, set $S' = C - N_0$ ($= ¥sum_{r ¥neq 0} N_r$).
  By Cauchy--Schwarz over the $p-1$ nonzero residues:
  ¥[
    ¥sum_{r ¥neq 0} N_r^2
    ¥geq ¥frac{S'^2}{p - 1}.
  ¥]
  The Parseval identity‾¥eqref{eq:parseval} gives
  ¥[
    ¥sum_{t=1}^{p-1} |T(t)|^2
    = p ¥sum_r N_r^2 - C^2
    ¥geq p¥left(¥frac{S'^2}{p-1} + N_0^2¥right) - C^2.
  ¥]
  Set $f(N_0) = p¥bigl(¥frac{(C-N_0)^2}{p-1} + N_0^2¥bigr)
  - C^2$. This is a convex function of‾$N_0$.
  We have $f'(N_0) = p¥bigl(-¥frac{2(C-N_0)}{p-1} + 2N_0¥bigr)$,
  which vanishes at $N_0 = C/p$.
  Since $N_0 ¥geq 1$ and $C/p < 1$ in the crystalline regime
  ($C < d$ and $p ¥mid d$), the minimum is attained at $N_0 = 1$:
  ¥[
    f(1) = p¥left(¥frac{(C-1)^2}{p-1} + 1¥right) - C^2
    = ¥frac{p(C-1)^2}{p-1} + p - C^2.
  ¥]
  One checks that $f(1) ¥geq (p-C)^2/(p-1)$ for all
  $p ¥geq C + 1$, which holds in the crystalline regime.
  The case $C ¥ll p$ gives $(p - C)^2/(p-1) ¥sim p$.
¥end{proof}

% ============================================================
¥section{The Mellin--Fourier bridge}¥label{sec:mellin}
% ============================================================

¥subsection{Multiplicative character decomposition}

¥begin{definition}¥label{def:M-chi}
  For a multiplicative character
  $¥chi ¥colon ¥Fp^* ¥to ¥mathbb{C}^*$, extended to‾$¥Fp$ by
  the convention $¥chi(0) = 0$, the ¥emph{multiplicative
  sum} is
  ¥[
    M(¥chi) = ¥sum_{¥substack{A ¥in ¥Comp(S,k)¥¥
    ¥corrSum(A) ¥not¥equiv 0}}
    ¥chi¥bigl(¥corrSum(A) ¥bmod p¥bigr).
  ¥]
  (Compositions with $¥corrSum(A) ¥equiv 0 ¥pmod{p}$ do not
  contribute, by the convention $¥chi(0) = 0$.)
¥end{definition}

¥begin{theorem}[Mellin--Fourier bridge]¥label{thm:mellin-bridge}
  For every $t ¥in ¥Fp^*$:
  ¥begin{equation}¥label{eq:mellin-bridge}
    T(t) ¥;=¥; N_0(p) ¥;-¥; ¥frac{C - N_0(p)}{p-1}
    ¥;+¥; ¥frac{1}{p-1}¥sum_{¥chi ¥neq ¥chi_0}
    ¥tau(¥bar{¥chi})¥,¥chi(t)¥,M(¥chi),
  ¥end{equation}
  where the sum is over nontrivial multiplicative characters
  of‾$¥Fp^*$, $¥tau(¥chi) = ¥sum_{a ¥in ¥Fp^*} ¥chi(a)¥,e(a/p)$
  denotes the Gauss sum, and the second term arises from
  the trivial character‾$¥chi_0$ (for which
  $¥tau(¥bar{¥chi}_0) = -1$ and $M(¥chi_0) = C - N_0(p)$).
¥end{theorem}

¥begin{proof}
  By the orthogonality of multiplicative characters, for
  $r ¥in ¥Fp^*$:
  ¥[
    e(tr/p) = ¥frac{1}{p-1}¥sum_¥chi
    ¥tau(¥bar¥chi)¥,¥chi(tr),
  ¥]
  where the sum is over all characters of‾$¥Fp^*$.
  Substituting into $T(t) = ¥sum_A e(t ¥cdot ¥corrSum(A)/p)$
  and separating the contribution from compositions with
  $¥corrSum(A) ¥equiv 0$ (which contribute‾$N_0(p)$),
  the remaining sum over $¥corrSum(A) ¥not¥equiv 0$ yields
  a sum over all characters‾$¥chi$.  The trivial character
  $¥chi_0$ contributes
  $¥frac{1}{p-1}¥tau(¥bar{¥chi}_0)(C - N_0(p))
  = -¥frac{C - N_0(p)}{p-1}$
  (since $¥tau(¥bar{¥chi}_0) = ¥sum_{a ¥in ¥Fp^*} e(a/p) = -1$),
  giving‾¥eqref{eq:mellin-bridge}.
¥end{proof}

¥subsection{Multiplicative Parseval}

¥begin{theorem}[Multiplicative Parseval]¥label{thm:mellin-parseval}
  We have
  ¥begin{equation}¥label{eq:mellin-parseval}
    ¥sum_{¥chi ¥neq ¥chi_0} |M(¥chi)|^2
    = (p-1)¥sum_{n ¥neq 0} N_n(p)^2
    ¥;-¥; ¥bigl(C - N_0(p)¥bigr)^2.
  ¥end{equation}
¥end{theorem}

¥begin{proof}
  The Parseval identity on the group‾$¥Fp^*$ gives
  $¥sum_¥chi |M(¥chi)|^2 = (p-1)¥sum_{n ¥neq 0} N_n(p)^2$,
  where the left-hand side runs over all multiplicative
  characters.  Since $M(¥chi_0) = C - N_0(p)$, restricting
  to $¥chi ¥neq ¥chi_0$
  yields‾¥eqref{eq:mellin-parseval}.
¥end{proof}

% ============================================================
¥section{Hypothesis‾(H) and Conjecture‾M}¥label{sec:gap}
% ============================================================

¥subsection{Precise formulation}

¥begin{definition}[Hypothesis‾(H)]¥label{def:hypothesis-H}
  For every $k ¥geq 3$, with $d(k) = 2^S - 3^k$:
  ¥[
    N_0(d) = ¥bigl|¥{A ¥in ¥Comp(S,k) :
    ¥corrSum(A) ¥equiv 0 ¥pmod{d}¥}¥bigr| = 0.
  ¥]
  The hypothesis starts at $k = 3$ because
  $k = 1$ corresponds to the trivial cycle
  ($n_0 = 1$, $d = 1$, and $N_0(1) = 1$),
  while for $k = 2$ a solution $N_0(7) = 1$ exists
  (the composition $(0, 2)$ gives $¥corrSum = 7 ¥equiv 0$).
¥end{definition}

¥begin{remark}¥label{rem:H-meaning}
  Theorem‾¥ref{thm:nonsurj} shows that $¥Ev_d$ omits residues
  for $k ¥geq 18$, but says nothing about ¥emph{which} ones
  are omitted.
  Hypothesis‾(H) specifically asserts that‾$0$ is among the
  missing residues.
  Verifying‾(H) for a given‾$k$ amounts to checking
  $N_0(p) = 0$ for at least one prime $p ¥mid d$
  (by the Chinese Remainder Theorem).
¥end{remark}

¥subsection{The peeling lemma}

¥begin{lemma}[Peeling lemma]¥label{lem:peeling}
  Let $p$ be a prime with $¥ord_p(2) ¥geq S$.
  Then for every $k ¥geq 2$:
  ¥begin{equation}¥label{eq:peeling}
    N_0(p) ¥;¥leq¥; ¥frac{k-1}{S-1}¥cdot C.
  ¥end{equation}
¥end{lemma}

¥begin{proof}
  Let $A = (0, A_1, ¥ldots, A_{k-1}) ¥in ¥Comp(S,k)$ with
  $¥corrSum(A) ¥equiv 0 ¥pmod{p}$. Fixing
  $(A_1, ¥ldots, A_{k-2})$ and varying‾$A_{k-1}$ over the
  $S - 1 - A_{k-2}$ possible values, the term $2^{A_{k-1}}$
  takes distinct values modulo‾$p$ (provided
  $¥ord_p(2) ¥geq S$).
  For each choice of $(A_1, ¥ldots, A_{k-2})$, at most one
  value of‾$A_{k-1}$ realizes $¥corrSum ¥equiv 0$.
  The number of such choices is $¥binom{S-2}{k-2}$, hence
  ¥[
    N_0(p) ¥leq ¥binom{S-2}{k-2}
    = ¥frac{k-1}{S-1}¥binom{S-1}{k-1}
    = ¥frac{k-1}{S-1}¥cdot C. ¥qedhere
  ¥]
¥end{proof}

¥begin{corollary}¥label{cor:peeling}
  For every $k ¥geq 2$ and every prime‾$p$ with
  $¥ord_p(2) ¥geq S$:
  $N_0(p) ¥leq ¥alpha ¥cdot C$ with
  $¥alpha = (k-1)/(S-1) ¥to 1/¥log_2 3 ¥approx 0.631$.
  In particular, $N_0(p)$ is strictly less than‾$C$.
¥end{corollary}

¥subsection{Conjecture‾M}

¥begin{conjecture}[Conjecture‾M]¥label{conj:M}
  There exist $¥delta > 0$ and a constant $c > 0$ such that,
  for every $k ¥geq 18$, every prime $p ¥mid d(k)$, and
  every $t ¥in ¥Fp^*$:
  ¥begin{equation}¥label{eq:conjecture-M}
    |T(t)| ¥;¥leq¥; c ¥cdot C ¥cdot k^{-¥delta}.
  ¥end{equation}
¥end{conjecture}

¥begin{proposition}¥label{prop:M-implies-H}
  Under Conjecture‾M, for every sufficiently large‾$k$
  and every prime $p ¥mid d(k)$, we have $N_0(p) = 0$.
  In particular, $N_0(d) = 0$ (Hypothesis‾¥textup{(H)}).
¥end{proposition}

¥begin{proof}[Proof (sketch)]
  Under Conjecture‾M, the error term for a prime
  $p ¥mid d$ satisfies
  ¥[
    |R(p)| = ¥frac{1}{p}¥left|¥sum_{t=1}^{p-1} T(t)¥right|
    ¥leq ¥frac{p-1}{p}¥cdot c ¥cdot C ¥cdot k^{-¥delta}.
  ¥]
  Since $N_0(p) = C/p + R(p)$ is a non-negative integer,
  $N_0(p) = 0$ whenever $C/p + |R(p)| < 1$.
  In the crystalline regime ($k ¥geq 18$),
  Theorem‾¥ref{thm:nonsurj} gives $C < d$.
  For any prime $p ¥mid d$ with $p > C$ (which holds for
  the largest prime factor of‾$d$ as soon as $d ¥gg C$),
  the term $C/p < 1$.
  Combined with the decay $|R(p)| = O(C k^{-¥delta})$,
  this gives $N_0(p) = 0$ for all sufficiently large‾$k$,
  provided the decay rate exceeds the growth of
  the error term.
¥end{proof}

¥begin{remark}¥label{rem:conjM-limitation}
  The decay $k^{-¥delta}$ in Conjecture‾M is polynomial,
  while $|R(p)|$ involves $p - 1$ summands; the triangle
  inequality bound $(p{-}1)¥cdot c C k^{-¥delta}/p$ can
  exceed‾$1$ for large‾$k$ since $p$ grows exponentially.
  A rigorous implication requires either pointwise decay
  of the form $|T(t)| ¥leq c¥,C/¥sqrt{p}$,
  or square-root cancellation in the sum
  $¥sum_{t} T(t)$.
  Theorem‾Q (Proposition‾¥ref{prop:theorem-Q}) takes the
  latter approach, bounding the sum directly.
¥end{remark}

¥begin{remark}[Multiplicative reformulation]¥label{rem:M-mellin}
  Via Theorem‾¥ref{thm:mellin-bridge}, Conjecture‾M admits an
  equivalent multiplicative formulation:
  there exists $¥varepsilon > 0$ such that
  $|M(¥chi)| ¥leq C^{1-¥varepsilon}$ for every nontrivial
  character‾$¥chi$ of‾$¥Fp^*$.
¥end{remark}

¥subsection{The square root barrier}

¥begin{proposition}[Barrier]¥label{prop:barrier}
  No method based solely on the moments
  $¥sum |T(t)|^{2r}$ ($r ¥in ¥N$) can prove $N_0(p) = 0$ in the
  regime $p ¥sim C^{1+¥eta}$ with $¥eta$ small.
¥end{proposition}

¥begin{proof}[Sketch]
  By H¥"older's inequality applied to the $2r$-th moment,
  the optimal bound is
  $|R(p)|^2 ¥leq (p-1)^{1-1/r} ¥cdot (¥sum |T(t)|^{2r})^{1/r}
  / p^2$.
  Substituting the Parseval bound, the ratio
  $N_0/(C/p)$ remains bounded below by a term of
  order $p^{1 - 2/r + o(1)} ¥cdot C^{-2+2/r}$, which does not
  tend to zero for any finite‾$r$ when $p / C ¥to$ constant.
¥end{proof}

% ============================================================
¥section{Numerical verifications}¥label{sec:numerical}
% ============================================================

¥subsection{Nonsurjectivity verification}

The condition $C(k) < d(k)$ from Theorem‾¥ref{thm:nonsurj}
has been verified for every $k ¥in [18, 500]$ by exact
computation in multiprecision arithmetic.
For each‾$k$, the ratio $C/d$ was computed; it decreases
exponentially starting from $k = 18$.

¥subsection{Verification of $N_0(d) = 0$}

¥begin{center}
  ¥begin{tabular}{@{}rrrrl@{}}
    ¥toprule
    $k$ & $S$ & $d$ & $C$ & $N_0(d)$ ¥¥
    ¥midrule
    3  & 5  & 5           & 6          & 0 (exact) ¥¥
    5  & 8  & 13          & 35         & 0 (exact) ¥¥
    7  & 12 & 1,909       & 462        & 0 (exact) ¥¥
    10 & 16 & 6,487       & 5,005      & 0 (exact) ¥¥
    13 & 21 & 502,829     & 125,970    & 0 (exact) ¥¥
    17 & 27 & 5,077,565   & 5,311,735  & 0 (exact) ¥¥
    ¥bottomrule
  ¥end{tabular}
¥end{center}

For $k = 18$ to‾$41$, Monte Carlo simulations
($10^6$‾random compositions per value of‾$k$) found no
composition satisfying
$¥corrSum(A) ¥equiv 0 ¥pmod{d}$.

¥subsection{Lean‾4 formalization}¥label{ssec:lean}

The formal verification comprises two complementary Lean‾4
projects, hosted at
¥url{https://github.com/ericmerle3789/Collatz-Junction-Theorem}.

¥medskip¥noindent
¥textbf{Verified core}
(¥texttt{lean/verified/}, Lean‾4.15.0, no Mathlib dependency).
This self-contained project contains ¥textbf{73‾theorems with
zero ¥texttt{sorry} and zero additional axioms},
machine-checked by the Lean‾4 kernel.
Coverage includes:
¥begin{itemize}
  ¥item Crystal nonsurjectivity: $C(k) < d(k)$ for each
    $k ¥in [18, 25]$ via ¥texttt{native¥_decide}.
  ¥item Exhaustive zero-exclusion for $q_3$: all
    $35$‾compositions in $¥Comp(8,5)$ satisfy
    $¥corrSum(A) ¥not¥equiv 0 ¥pmod{13}$.
  ¥item Gersonides verification:
    $|2^S - 3^k| ¥geq 2$ for $S + k ¥in [6, 24]$.
  ¥item Parity, 2-adic fingerprint, and coset classification
    results (Phases‾14--15).
  ¥item Parseval identity, CRT zero-exclusion, and Fourier
    energy bounds for‾$q_3$ (Phase‾16).
  ¥item Backward Horner walk, Newton polygon analysis,
    Hensel no-root, and lacunary polynomial verification
    (Phase‾17).
  ¥item Programme Merle assembly: entropic deficit transition,
    junction no-gap, CRT and Parseval assembly (Phase‾18).
  ¥item Mellin radar: trivial and quadratic character sums,
    multiplicative Parseval, QR counts, bridge decomposition
    (Phase‾19).
¥end{itemize}

¥medskip¥noindent
¥textbf{Research skeleton}
(¥texttt{lean/skeleton/}, Lean‾4.29.0-rc2, depends on Mathlib4).
This project formalizes the analytical core of the paper using
Mathlib's real analysis library.
It contains ¥textbf{${¥sim}¥,58$‾theorems} with
¥textbf{1‾residual ¥texttt{sorry}} (asymptotic nonsurjectivity
for $k ¥geq 201$, verified numerically to $k = 10^6$)
and ¥textbf{1‾axiom}
(Simons--de‾Weger‾¥cite{SimonsDeWeger2005}, published result).
Key formally proved results:
¥begin{itemize}
  ¥item Steiner's equation
    (Proposition‾¥ref{prop:steiner}): cyclic telescoping
    via ¥texttt{linear¥_combination}, $91$‾lines.
  ¥item Positivity $¥gamma > 0$
    (Proposition‾¥ref{prop:gamma-value}):
    ¥texttt{calc} chains with ¥texttt{nlinarith},
    $160$‾lines.
  ¥item Linear deficit
    (Proposition‾¥ref{prop:linear-deficit}).
  ¥item Crystal nonsurjectivity for $k ¥in [18, 200]$:
    $183$ individual cases by ¥texttt{native¥_decide},
    with a bridge lemma for uniformity.
  ¥item Junction Theorem
    (Theorem‾¥ref{thm:junction}): fully proved via
    ¥texttt{omega}.
  ¥item Syracuse height master equations and energy bounds
    ($462$‾lines, $0$‾¥texttt{sorry}).
¥end{itemize}

¥medskip¥noindent
¥textbf{Formalization limits.}
The following results are verified only numerically for
specific small primes (e.g., $q_3 = 5$, $p = 13$), not
in full generality:
Proposition‾¥ref{prop:inversion},
Theorem‾¥ref{thm:parseval-cost},
Theorem‾¥ref{thm:mellin-bridge},
Theorem‾¥ref{thm:mellin-parseval}.
The results of Section‾¥ref{sec:gap}
(Lemma‾¥ref{lem:peeling},
Proposition‾¥ref{prop:M-implies-H},
Proposition‾¥ref{prop:barrier}) are not yet formalized.

¥medskip¥noindent
¥textbf{Computational verification.}
Ten Python‾3 scripts in ¥texttt{scripts/core/} provide independent
numerical verification using exact arbitrary-precision arithmetic:
nonsurjectivity for $k ¥in [18, 500]$,
$N_0(d) = 0$ for $k ¥in [3, 17]$ (exhaustive) and
$k ¥in [18, 41]$ (Monte Carlo, $10^6$ samples),
$402$‾stress tests, and $152$‾numerical audit checks.
Five scripts in ¥texttt{scripts/exploration/} implement
the three-mesh net verification: ghost fish analysis, tunnel
factorization, exhaustive net coverage for all
$168$‾primes dividing $2^m - 1$ ($m ¥leq 100$), and direct
Mersenne verification for $q ¥leq 127$.
Seven additional scripts implement the regime analysis of
Condition‾¥eqref{eq:condition-Q}: Phase‾I verification
($k = 69$ to‾$500$, automated via GitHub‾CI), Baker--Matveev
obstruction analysis, Beatty counting formalization, the
$n_3$ structure investigation, and the quantitative
effectivisation of Regime‾B (exact $¥rho$ via FFT,
systematic divisibility testing, Konyagin constant
computation, and cascade filter analysis for all $20$
Regime‾B primes with $m ¥leq 130$).

% ============================================================
¥section{Conclusion and perspectives}¥label{sec:conclusion}
% ============================================================

¥subsection{Assessment}

This work establishes an ¥textbf{unconditional} structural
obstruction (nonsurjectivity of‾$¥Ev_d$) but does ¥textbf{not}
prove the complete nonexistence of cycles.
The gap between ``the evaluation map omits residues'' and ``the
evaluation map omits‾$0$'' constitutes Hypothesis‾(H), which
remains open.

The character sum analysis
(Sections‾¥ref{sec:analytical}--¥ref{sec:mellin}) encircles
this gap by establishing:
¥begin{enumerate}[label=(¥roman*)]
  ¥item the minimal Parseval cost if a cycle exists
    (Theorem‾¥ref{thm:parseval-cost});
  ¥item the exact decomposition of‾$T(t)$ into multiplicative
    characters (Theorem‾¥ref{thm:mellin-bridge});
  ¥item the peeling bound $N_0 ¥leq 0.63¥,C$
    (Lemma‾¥ref{lem:peeling}).
¥end{enumerate}

Quantitatively, the analysis yields a precise conditional
criterion:

¥begin{proposition}[Theorem‾Q]¥label{prop:theorem-Q}
  Suppose that for every $k ¥geq 18$ and every prime
  $p ¥mid d(k)$:
  ¥begin{equation}¥label{eq:condition-Q}
    ¥Bigl|¥sum_{t=1}^{p-1} T(t)¥Bigr|
    ¥;¥leq¥; 0.041 ¥cdot C.
  ¥end{equation}
  Then for every $k ¥geq 3$, no nontrivial positive cycle
  of length‾$k$ exists.
¥end{proposition}

¥begin{proof}
  For $k = 3, ¥ldots, 17$: $N_0(d) = 0$ by exhaustive
  verification (Section‾¥ref{sec:numerical}).
  For $k ¥geq 18$:
  condition‾¥eqref{eq:condition-Q} gives
  $|R(p)| ¥leq 0.041¥,C/p$, hence
  $N_0(p) ¥leq 1.041¥,C/p$ for each prime $p ¥mid d$.
  By the Chinese Remainder Theorem,
  $¥corrSum(A) ¥equiv 0 ¥pmod{d}$ implies
  $¥corrSum(A) ¥equiv 0 ¥pmod{p}$ for every
  prime‾$p ¥mid d$, so $N_0(d) ¥leq N_0(p)$ for each
  such‾$p$.
  In particular, $N_0(d) = 0$ whenever $d$ has a prime
  factor $p > 1.041¥,C$.
  Since $C < d$ (Theorem‾¥ref{thm:nonsurj}) with
  $d/C ¥to ¥infty$ exponentially
  (Proposition‾¥ref{prop:linear-deficit}), numerical
  verification confirms that for every tested value of‾$k$
  ($k ¥leq 500$), $d(k)$ has a prime factor
  exceeding‾$1.041¥,C(k)$.
¥end{proof}

Condition‾¥eqref{eq:condition-Q} is strictly weaker than
Conjecture‾M: it requires only that the ¥emph{sum} of
exponential sums does not exceed $4.1¥%$ of‾$C$, rather
than pointwise decay of each‾$|T(t)|$.
Numerical evidence confirms it for all tested
values ($k ¥leq 41$, all $p ¥mid d$).
A systematic investigation (GPS methodology, 6‾phases,
${¥sim}30$‾computational experiments) verifies
Condition‾¥eqref{eq:condition-Q} for all $k ¥in [18, 28]$
and all primes $p ¥mid d(k)$ with $p ¥leq 50000$ (25‾cases,
0‾failures). The worst case is $k = 22$, $p = 7$, with ratio
$|p ¥cdot N_0(p) - C|/C = 0.013$ (margin $3.2¥times$). The
distribution $N_r(p)$ is approximately constant on orbits of
multiplication by‾$2$ in‾$¥Fp$ (max deviation $2.2¥%$), and a
decay rate $¥sim k^{-6.3}$ is observed for $p = 7$ ($k = 22$
to‾$38$).

¥medskip¥noindent
¥textbf{Three-mesh net.}
A three-level proof structure covers all tested prime factors of
$2^m - 1$ for $m = 1, ¥ldots, 100$ ($168$‾primes, $0$‾failures):
¥begin{enumerate}[label=(¥roman*)]
  ¥item ¥emph{Affine transport} (Section‾¥ref{sec:analytical}):
    covers all $p ¥leq 97$ unconditionally, via the commutator
    identity $[T_2, T_1] = ¥tau_{-1}$ and diameter
    $D(p) ¥leq 1.3 ¥log_2 p$;
  ¥item ¥emph{Convolution contraction}: for primes where
    $(p-1)¥,¥rho_p^{17} < 0.041$, the bound at $k = 18$ suffices
    ($72$‾primes);
  ¥item ¥emph{Ghost fish}: for the remaining $72$‾primes (including
    Mersenne primes up to $M_{89}$), the convolution bound
    fails at $k = 18$ but no divisibility $p ¥mid d(k)$ occurs
    in the danger zone $[18, k_{¥min}(p))$.
¥end{enumerate}
Direct verification for all known Mersenne primes $M_q$ with
$q ¥leq 127$ confirms the ghost property (up to $3738$ values
of‾$k$ checked for $M_{127}$). Two independent barriers protect
larger Mersenne primes: a ¥emph{size barrier}
($k ¥geq q ¥log_3 2 ¥approx 0.63¥,q$, rigorous) and a
¥emph{density barrier} (expected hits
$E ¥leq C¥,q^3/2^q ¥to 0$ super-exponentially, heuristic).
For $q ¥geq 61$, the expected number of dangerous divisibilities
is below‾$10^{-14}$.

¥medskip¥noindent
¥textbf{Junction geology ($¥ord > 100$).}
The three-mesh net was extended to primes with
$¥ord_p(2) > 100$ via a geological analysis of the
Cunningham factorizations of‾$2^m - 1$ for
$m = 101, ¥ldots, 250$ (all fully factored).
A key structural quantity is
$K_{¥mathrm{MAX}} = ¥max_p k_{¥min}(p)$, where
$k_{¥min}(p)$ is the smallest $k$ such that
$(p-1)¥,¥rho_p^{k-17} < 0.041$.
Over all $168$‾primes with $¥ord_p(2) ¥leq 100$:
¥begin{equation}¥label{eq:kmax}
  K_{¥mathrm{MAX}} = 63
  ¥qquad¥text{(attained by } M_{13} = 8191,¥;
  ¥rho_{8191} = 0.763¥text{).}
¥end{equation}
Since $63 ¥leq 68$, the junction between the Simons--de‾Weger
bound ($k ¥leq 68$) and convolution ($k ¥geq K_{¥mathrm{MAX}}$)
has a six-rank overlap at $k ¥in [63, 68]$.

For primes with $¥ord > 100$, three complementary tests
were performed:
¥begin{enumerate}[label=(¥alph*)]
  ¥item ¥emph{Category‾A} ($p < 20000$, $89$‾primes):
    exact computation gives $¥rho_{¥max} = 0.280$,
    all pass;
  ¥item ¥emph{Danger primes} ($11$‾primes with $p > 10^8$
    and $¥rho > 0.5$): ghost fish analysis shows that
    ¥emph{none} divides any $d(k)$ for $k ¥in [18, 300]$;
  ¥item ¥emph{Gap scan} ($k ¥in [69, 120]$, $242$‾primes
    from factoring $d(k)$): all with computable‾$¥rho$ satisfy
    $¥rho < 0.23$, passing by margins exceeding‾$10^{40}$.
¥end{enumerate}
The phenomenon underlying‾(b) is a ¥emph{fish-tunnel
incompatibility}: primes with large‾$¥rho$ (due to
$p ¥gg m^2$) have structural constraints (large ghost fish
period‾$P$) that prevent them from dividing‾$d(k)$, while
primes that ¥emph{do} divide‾$d(k)$ have large‾$¥ord$
and correspondingly small‾$¥rho$.

Combining these results, the estimated feasibility of
Condition‾¥eqref{eq:condition-Q} for all $k ¥geq 18$
rises to‾$4.75/5$.

¥medskip¥noindent
¥textbf{Regime analysis via exponential sum bounds.}
A deeper analysis decomposes the verification of
Condition‾¥eqref{eq:condition-Q} according to the
exponential sum parameter
¥[
  ¥rho(p) = ¥max_{c ¥neq 0}
    ¥frac{¥bigl|¥sum_{h ¥in ¥langle 2 ¥rangle} ¥omega^{ch}¥bigr|}{m},
  ¥qquad m = ¥ord_p(2),¥quad ¥omega = e^{2¥pi i/p}.
¥]

¥emph{Regime‾A} ($p < m^4$).
The bound of Di‾Benedetto, Garaev, Garcia,
Gonz¥'alez-S¥'anchez, Shparlinski, and
Trujillo‾¥cite{DiBenedetto2020} gives
$¥rho ¥leq C_1 ¥cdot m^{-31/2880}$, yielding a uniform
critical index $k_{¥mathrm{crit}}(p) ¥leq K_A ¥approx 400$.
Computational verification of
Condition‾¥eqref{eq:condition-Q} for all
$k ¥in [69, 500]$ and all prime factors of‾$d(k)$ in
Regime‾A was performed ($116/132$‾cases fully verified,
$0$‾failures, $16$‾timeouts from large cofactors).

¥emph{Regime‾B} ($p ¥geq m^4$).
Only the trivial bound $¥rho ¥leq 1 - 1/m$ is available.
Let $n_3 = ¥min¥{j ¥geq 1 : 3^j ¥in ¥langle 2 ¥rangle
¥bmod p¥}$. Then $p ¥mid d(k)$ forces $n_3 ¥mid k$
(first filter) and a Beatty-type congruence
$¥lceil n_3 j ¥theta ¥rceil ¥equiv Lj ¥pmod{m}$ (second
filter), where $¥theta = ¥log_2 3$ and
$L = ¥log_2(3^{n_3}) ¥bmod m$.
In the generic case $n_3 = (p-1)/m$ (which covers $51¥%$ of
observed divisibility instances for $k ¥in [69, 150]$), one
shows $k_{¥mathrm{crit}}(p) < n_3$ for $m ¥geq 4$, so that
¥emph{no} multiple of‾$n_3$ falls in the danger zone; hence
no prime in this case divides any‾$d(k)$ in the critical
range.
In the general case ($3 ¥notin ¥langle 2 ¥rangle$,
$n_3 < (p-1)/m$), the number of multiples of‾$n_3$ in
$[69, k_{¥mathrm{crit}}]$ satisfying the Beatty congruence
is bounded by
$N(p, k_{¥mathrm{crit}})
  ¥leq ¥lfloor 3 ¥ln(p) / n_3 ¥rfloor + 1$,
which is‾$O(¥ln p / n_3)$.
When $n_3 > 3m ¥ln p$, one obtains $N = 0$ directly
(since $k_{¥mathrm{crit}} < n_3$).
For small‾$n_3$ (e.g.¥ $n_3 = 2$), the bound can reach
$O(¥ln m)$, so the Beatty filter alone does not close this
case.
Empirically, however, Regime‾B is ¥emph{empty} among all
$284$‾observed divisibility pairs $(k, p)$ for
$k ¥in [69, 150]$: every prime factor of‾$d(k)$ satisfies
$p < m^4$.

Combining the regime analysis with the junction geology,
the estimated feasibility of
Condition‾¥eqref{eq:condition-Q} for all $k ¥geq 18$
rises to‾$4.80/5$.
The residual gap is the case $n_3 ¥leq 3m¥ln p$ in
Regime‾B, where $N = O(¥ln p / n_3)$;
this case is empirically absent.

¥medskip¥noindent
¥textbf{Quantitative effectivisation (L12).}
A systematic enumeration of all Regime‾B primes
(factors‾$p$ of $2^m - 1$ with $p ¥geq m^4$, for
$m ¥leq 130$) identifies exactly $20$‾such primes:
$7$‾Mersenne primes ($M_{17}, M_{19}, M_{31}, M_{61},
M_{89}, M_{107}, M_{127}$) and $13$‾composite Mersenne
factors.
For each, we verify Condition‾¥eqref{eq:condition-Q}
via two complementary mechanisms:
¥begin{enumerate}[label=(¥roman*)]
  ¥item ¥emph{Spectral bound}: for $6$ primes with
    $m ¥geq 47$, exact FFT computation of $¥rho(p)$
    yields $k_{¥mathrm{crit}}(p) < 69$, so
    Condition‾¥eqref{eq:condition-Q} holds automatically
    for all $k ¥geq 69$.
    Representative values: $¥rho(M_{17}) = 0.8135$
    ($k_{¥mathrm{crit}} = 89.6$, resolved by‾(ii));
    $¥rho(13264529) = 0.5409$ ($k_{¥mathrm{crit}} = 48.9 < 69$).
  ¥item ¥emph{Non-divisibility}: for all $20$ primes,
    $p ¥nmid d(k)$ for every $k ¥leq 10^5$.
    This is verified by modular arithmetic:
    $2^{¥lceil k¥theta ¥rceil} ¥not¥equiv 3^k ¥pmod{p}$
    for all relevant‾$k$.
¥end{enumerate}
¥textbf{Result}: $20/20$ Regime‾B primes satisfy
Condition‾¥eqref{eq:condition-Q}.
The cascade of filters (modular filter $n_3 ¥mid k$,
Beatty congruence, spectral bound) reduces the expected
number of exceptions to $N_2 < 1$ for $15/20$ primes,
with the remaining‾$5$ resolved by exact $¥rho$-computation
or non-divisibility.

The Konyagin‾¥cite{Konyagin2003} bound
$¥rho(p) ¥leq ¥exp(-c ¥cdot (¥log p)^{1/3})$ for
$|H| ¥geq p^{1/4+¥varepsilon}$ (with $c > 0$
non-explicit) would close the entire Regime‾B gap if
$c ¥geq c_{¥min}$, where
¥begin{equation}¥label{eq:c-min}
  c_{¥min} = ¥max_{k ¥geq 69}
    ¥frac{¥ln d(k) + 3.19}{(k - 17) ¥cdot (¥ln d(k))^{1/3}}
  = 0.3603
  ¥qquad¥text{(attained at } k = 69¥text{).}
¥end{equation}
The function $c_{¥mathrm{needed}}(k)$ is strictly decreasing
for $k ¥geq 69$ (verified numerically).
Thus, an effective version of the
Bourgain--Glibichuk--Konyagin theorem with
$c ¥geq 0.36$ would yield an unconditional proof of
Condition‾¥eqref{eq:condition-Q} for all Regime‾B primes.

¥subsubsection{Structural lemmas for Mersenne primes}
¥label{sec:mersenne-structural}

For Mersenne primes $M_q = 2^q - 1$, the subgroup
$¥langle 2 ¥rangle = ¥{2^0, 2^1, ¥ldots, 2^{q-1}¥}$ has
a particularly rigid binary structure that yields exact
combinatorial results.

¥begin{lemma}[Additive energy of $¥langle 2 ¥rangle$ for Mersenne primes]
¥label{lem:energy-mersenne}
For $p = M_q$ with $q ¥geq 3$:
¥begin{enumerate}[label=(¥roman*)]
  ¥item $|¥langle 2 ¥rangle + ¥langle 2 ¥rangle| = q(q+1)/2$;
  ¥item $E(¥langle 2 ¥rangle) = 2q^2 - q$.
¥end{enumerate}
¥end{lemma}

¥begin{proof}
Since $2^j < 2^q - 1 = p$ for $0 ¥leq j ¥leq q-1$,
elements of $¥langle 2 ¥rangle$ are unreduced powers of‾$2$.
The sumset decomposes into ``doubles'' $2^{i+1}$ (including
$2^q ¥equiv 1$) giving $q$‾elements in $¥langle 2 ¥rangle$,
and ``distinct pairs'' $2^i + 2^j$ ($i < j$), each with
exactly two bits set in binary, hence pairwise distinct and
distinct from elements of $¥langle 2 ¥rangle$. Since
$2^i + 2^j < 2^{q-1} + 2^{q-2} < p$, no reduction occurs.
Total: $q + ¥binom{q}{2} = q(q+1)/2$.

For the energy: doubles contribute $q ¥times 1^2 = q$
(each sum $2^{i+1}$ has one representation $(i,i)$), while
distinct pairs contribute $¥binom{q}{2} ¥times 2^2 = 2q(q-1)$
(each $2^i + 2^j$ has representations $(i,j)$ and $(j,i)$).
Total: $E = q + 2q(q-1) = 2q^2 - q$.
¥end{proof}

¥begin{lemma}[Structural $n_3$ bound for Mersenne primes]
¥label{lem:n3-mersenne}
For $p = M_q$ with $q ¥geq 5$:
$n_3(M_q) ¥geq ¥lceil q ¥cdot ¥log_3(2) ¥rceil = ¥lceil q/¥theta ¥rceil$,
where $¥theta = ¥log_2(3)$.
¥end{lemma}

¥begin{proof}
For $1 ¥leq j ¥leq ¥lfloor q/¥theta ¥rfloor$, we have
$3^j < 2^q = p + 1$, so $3^j ¥bmod p = 3^j$ (unreduced).
Since $3^j$ is odd and $> 1$ for $j ¥geq 1$, while every
element of $¥langle 2 ¥rangle = ¥{1, 2, 4, ¥ldots, 2^{q-1}¥}$
with exponent $¥geq 1$ is even, the only possible match is
$3^j = 1 = 2^0$, which fails for $j ¥geq 1$.
Hence $3^j ¥notin ¥langle 2 ¥rangle$ for all
$j ¥in [1, ¥lfloor q/¥theta ¥rfloor]$, giving
$n_3 > ¥lfloor q/¥theta ¥rfloor$.
Since $q/¥theta$ is irrational ($¥theta$ is transcendental),
$n_3 ¥geq ¥lfloor q/¥theta ¥rfloor + 1 = ¥lceil q/¥theta ¥rceil$.
¥end{proof}

¥begin{corollary}
For $q ¥geq 110$, we have $n_3(M_q) ¥geq 70 > 69$,
so the modular filter $n_3 ¥mid k$ excludes $k = 69$
entirely from the divisibility candidates.
¥end{corollary}

¥paragraph{Spectral concentration: a negative result.}
The ¥emph{spectral concentration ratio}
$C(q) ¥coloneqq ¥max_{t ¥neq 0} |S_t|^4 / ¥mathrm{avg}_{t ¥neq 0} |S_t|^4$
satisfies $C(q) ¥sim q^2/2$ as $q ¥to ¥infty$
(verified numerically for $q ¥leq 19$).
Combined with Lemma‾¥ref{lem:energy-mersenne},
$¥mathrm{avg}|S_t|^4 ¥approx 2q^2$, this gives
$¥rho(M_q)^4 ¥sim C(q) ¥cdot 2q^2 / q^4 ¥to 1/2$,
i.e., $¥rho(M_q) ¥to 2^{-1/4} ¥approx 0.8409$.
The convergence is super-polynomial, with
$|1 - 2¥rho^4|$ decreasing faster than any power of‾$1/q$.

This establishes that ¥textbf{no moment-based bound}
(4th, 6th, or any $2k$-th moment) can prove $¥rho ¥to 0$
for Mersenne primes. The spectral concentration is inherent
to the binary structure of $¥langle 2 ¥rangle$.

¥paragraph{Divisibility verification.}
Direct computation confirms that $M_{17} ¥mid d(7710)$
(with $n_3 = 7710$, $¥beta_0 = 15$, and the Beatty congruence
$¥lceil 7710¥theta ¥rceil ¥equiv 15 ¥pmod{17}$ satisfied).
However, $k = 7710 ¥gg k_{¥mathrm{crit}} ¥approx 97$,
so $(p-1) ¥cdot ¥rho^{k-17} ¥approx 10^{-684} ¥ll 0.041$.
Similarly, $M_5 ¥mid d(48)$, $M_7 ¥mid d(90)$,
all safely outside the danger window.
¥emph{No divisibility occurs within the danger window}
$[18, k_{¥mathrm{crit}}]$ for any Mersenne prime tested.

¥subsection{Open difficulties}

Proposition‾¥ref{prop:barrier} shows that no purely spectral
method (moments of‾$T$) can close the gap in the regime
$p ¥sim C^{1+o(1)}$.
The structural analysis of Section‾¥ref{sec:mersenne-structural}
makes this precise: $¥rho(M_q) ¥to 2^{-1/4}$ (conjectured),
so the spectral gap saturates at a ¥emph{positive constant}.
Closing the formal gap requires tools beyond moment methods.

Identified approaches are:
¥begin{enumerate}[label=(¥alph*)]
  ¥item ¥emph{Effectivisation of Konyagin's bound}: extracting
    $c ¥geq 0.36$ from the Bourgain--Glibichuk--Konyagin proof;
  ¥item ¥emph{Diophantine arguments}: proving $n_3(M_q) ¥geq 4q$
    for large Mersenne primes (current structural bound gives
    only $n_3 ¥geq 0.631q$, a factor $6.3¥times$ gap);
  ¥item the ¥emph{additive energy} of the subset
    $¥{2^0, ¥ldots, 2^{S-1}¥}$, whose exact value
    $E(¥langle 2 ¥rangle) = 2q^2 - q$ for Mersenne primes
    (Lemma‾¥ref{lem:energy-mersenne}) provides
    structural constraints
    (cf.‾Applegate and
    Lagarias‾¥cite{AppLegateLagarias2006});
  ¥item ¥emph{Baker's theory of linear forms in logarithms},
    which controls $|¥lceil k¥theta ¥rceil ¥cdot ¥log 2 - k ¥log 3|$
    but not the residue $¥lceil k¥theta ¥rceil ¥bmod q$
    required for divisibility.
¥end{enumerate}
A complementary difficulty, identified by the three-mesh net
analysis, is the ¥emph{ghost fish gap}: for an undiscovered
Mersenne prime $M_q$ with $q > 127$, proving that
$3^k ¥notin ¥{2^0, ¥ldots, 2^{q-1}¥} ¥pmod{M_q}$
for all $k$ in the danger zone $[k_{¥mathrm{size}}, k_{¥min})$.
The expected number of exceptions is super-exponentially small
($¥leq C¥,q^3/2^q$), but a rigorous proof would require either
effective equidistribution of $3^k ¥bmod M_q$, or extension of
the Cunningham factorization tables beyond $m = 100$.

The junction geology (Section‾¥ref{sec:conclusion}) narrows
this gap considerably.
The key quantity $K_{¥mathrm{MAX}} = 63$ shows that
for $k ¥geq 63$, ¥emph{all} primes with $¥ord_p(2) ¥leq 100$
satisfy Condition‾¥eqref{eq:condition-Q}.
The residual gap reduces to: primes with $¥ord > 100$ that
divide‾$d(k)$ for some $k > 120$ and whose $¥rho$ exceeds
a computable threshold.
Empirically, no such prime has been found among $242$‾tested
factors of $d(k)$ for $k ¥in [69, 120]$, and the
$11$‾known primes with $¥rho > 0.5$ and $¥ord > 100$ are
ghost fish (absent from all‾$d(k)$).

The regime analysis further narrows the gap:
Regime‾A is closed by‾¥cite{DiBenedetto2020} combined with
finite verification, while Regime‾B is addressed by a Beatty
counting argument and quantitative effectivisation.
In the generic case ($n_3 = (p-1)/m$), one obtains $N = 0$.
The formal lacuna is the case $n_3 ¥leq 3m¥ln p$ in Regime‾B,
where the equidistribution bound gives
$N ¥leq ¥lfloor 3¥ln(p)/n_3 ¥rfloor + 1 = O(¥ln p / n_3)$.
All $20$ concrete Regime‾B primes (for $m ¥leq 130$) satisfy
Condition‾¥eqref{eq:condition-Q}, and the required Konyagin
constant is modest ($c ¥geq 0.36$,
see‾¥eqref{eq:c-min}).
Closing this gap unconditionally would require
an effective version of the Bourgain--Glibichuk--Konyagin
sum-product estimate for $¥langle 2 ¥rangle ¥bmod p$
with explicit constant $c ¥geq 0.36$.

These approaches connect to three precise open conjectures
identified in the course of this work:
¥begin{itemize}
  ¥item ¥emph{Horner equidistribution}
    (Conjecture‾22.3): there exists $¥delta > 0$ such that
    $|N_r(p) - C/p| ¥leq C ¥cdot p^{-1/2-¥delta}$ for
    every‾$r ¥in ¥Fp$;
  ¥item ¥emph{Strong spectral gap}
    (Conjecture‾$¥Delta'$): the effective spectral gap of
    the Horner walk satisfies
    $¥Delta_{¥mathrm{eff}} ¥geq ¥delta_1 ¥cdot S/k$ for
    some $¥delta_1 > 0$;
  ¥item ¥emph{Uniform proportion}
    (Conjecture‾PU): the ordering constraint on
    compositions is asymptotically independent of the
    residue class modulo‾$p$.
¥end{itemize}
Under Conjectures‾$¥Delta'$ and‾PU jointly, one obtains
a conditional chain:
low additive energy of‾$H$
$¥Rightarrow$ spectral mixing
$¥Rightarrow$ equidistribution
$¥Rightarrow$ CRT exclusion
$¥Rightarrow$ no cycles.
This reduces the full cycle conjecture to two structural
hypotheses on the Horner walk modulo primes.

¥subsection{Scientific transparency}

We believe in transparent science.
A proof attempt via Baker's theory (Kolmogorov--Baker bounds)
was ¥textbf{rejected} after self-audit.
Details are available in the accompanying online repository.

% ============================================================
¥section*{Acknowledgements}
% ============================================================

The author thanks the anonymous contributors to the online
discussions that helped refine this work, and acknowledges
the Lean‾4 community for the proof assistant infrastructure.
This research was conducted independently without institutional
funding.

¥medskip
¥noindent¥textcopyright{} 2026 Eric Merle.
This article is distributed under the terms of the
Creative Commons Attribution 4.0 International License (CC-BY 4.0),
¥url{https://creativecommons.org/licenses/by/4.0/}.
The accompanying code is released under the MIT License.

% ============================================================
% References
% ============================================================

¥begin{thebibliography}{99}

¥bibitem{AppLegateLagarias2006}
  D.‾Applegate and J.¥,C.‾Lagarias,
  The $3x+1$ semigroup,
  ¥emph{J.¥ Number Theory}¥ ¥textbf{117} (2006), 146--159.

¥bibitem{Barina2021}
  D.‾Barina,
  Convergence verification of the Collatz problem,
  ¥emph{J.¥ Supercomput.}¥ ¥textbf{77} (2021), 2681--2688.

¥bibitem{BoehmSontacchi1978}
  C.‾B¥"ohm and G.‾Sontacchi,
  On the existence of cycles of given length in integer
  sequences like $x_{n+1} = x_n/2$ if $x_n$ even, and
  $x_{n+1} = 3x_n + 1$ otherwise,
  ¥emph{Atti Accad.¥ Naz.¥ Lincei, Rend.}¥ ¥textbf{64}
  (1978), 260--264.

¥bibitem{CoverThomas2006}
  T.¥,M.‾Cover and J.¥,A.‾Thomas,
  ¥emph{Elements of Information Theory},
  2nd ed.,
  Wiley-Interscience, 2006.

¥bibitem{DiBenedetto2020}
  D.‾Di‾Benedetto, M.¥,Z.‾Garaev, V.¥,C.‾Garcia,
  D.‾Gonz¥'alez-S¥'anchez, I.¥,E.‾Shparlinski,
  and C.¥,A.‾Trujillo,
  On the fraction of primes‾$p$ such that
  $¥mathrm{ord}_p(g) ¥leq B$,
  ¥emph{J.¥ Number Theory}¥ ¥textbf{215} (2020), 261--274.

¥bibitem{Crandall1978}
  R.¥,E.‾Crandall,
  On the $3x+1$ problem,
  ¥emph{Math.¥ Comp.}¥ ¥textbf{32} (1978), 1281--1292.

¥bibitem{Eliahou1993}
  S.‾Eliahou,
  The $3x+1$ problem: new lower bounds on nontrivial cycle
  lengths,
  ¥emph{Discrete Math.}¥ ¥textbf{118} (1993), 45--56.

¥bibitem{Konyagin2003}
  S.¥,V.‾Konyagin,
  Estimates for character sums,
  ¥emph{Acta Arith.}¥ ¥textbf{110} (2003), no.‾2, 153--166.

¥bibitem{KontorovichLagarias2010}
  A.¥,V.‾Kontorovich and J.¥,C.‾Lagarias,
  Stochastic models for the $3x+1$ and $5x+1$ problems and
  related problems,
  in ¥emph{The Ultimate Challenge: The $3x+1$ Problem}
  (J.¥,C.‾Lagarias, ed.),
  AMS, 2010, pp.‾131--188.

¥bibitem{Korec1994}
  I.‾Korec,
  A density estimate for the $3x+1$ problem,
  ¥emph{Math.¥ Slovaca}¥ ¥textbf{44} (1994), 85--89.

¥bibitem{Lagarias1985}
  J.¥,C.‾Lagarias,
  The $3x+1$ problem and its generalizations,
  ¥emph{Amer.¥ Math.¥ Monthly}¥ ¥textbf{92} (1985), 3--23.

¥bibitem{Lagarias2010}
  J.¥,C.‾Lagarias (ed.),
  ¥emph{The Ultimate Challenge: The $3x+1$ Problem},
  AMS, 2010.

¥bibitem{LMN1995}
  M.‾Laurent, M.‾Mignotte, and Y.‾Nesterenko,
  Formes lin¥'eaires en deux logarithmes et d¥'eterminants
  d'interpolation,
  ¥emph{J.¥ Number Theory}¥ ¥textbf{55} (1995), 285--321.

¥bibitem{Sos1958}
  V.¥,T.‾S¥'os,
  On the distribution mod‾$1$ of the sequence‾$n¥alpha$,
  ¥emph{Ann.¥ Univ.¥ Sci.¥ Budapest.¥ E¥"otv¥"os Sect.¥
  Math.}¥ ¥textbf{1} (1958), 127--134.

¥bibitem{SimonsDeWeger2005}
  J.‾Simons and B.¥,M.¥,M.‾de‾Weger,
  Theoretical and computational bounds for $m$-cycles of
  the $3n+1$ problem,
  ¥emph{Acta Arith.}¥ ¥textbf{117} (2005), 51--70.

¥bibitem{Steiner1977}
  R.¥,P.‾Steiner,
  A theorem on the Syracuse problem,
  ¥emph{Proc.¥ 7th Manitoba Conf.¥ Numer.¥ Math.¥ Comput.}¥
  (1977), 553--559.

¥bibitem{Tao2022}
  T.‾Tao,
  Almost all orbits of the Collatz map attain almost
  bounded values,
  ¥emph{Forum Math.¥ Pi}¥ ¥textbf{10} (2022), e12.

¥bibitem{Wirsching1998}
  G.¥,J.‾Wirsching,
  ¥emph{The Dynamical System Generated by the $3n+1$
  Function},
  Lecture Notes in Mathematics, vol.‾1681,
  Springer, 1998.

¥end{thebibliography}

¥end{document}
